How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-03 tse-paper/abstract.tex
small corrections introduction, Closes #20

Build logs are a textual by-product that software build processes
create, often as part of their Continuous Integration (CI)
pipeline. Build logs can serve as a paramount source of
information for developers to understand and debugg build or
test failures. Manually
extracting the important chunks of information, though, is akin to
finding a needle in a haystack. Recently, researchers and practitioners have begun attempts to partly automate
this time-consuming activity, with the proposition to not only ease developers' tasks but to enable a new class of automated on-ward processing applications.
In this paper, we want to give a systematic overview of the state of the art of the emerging field of build log analysis. To this end, we first survey and categorize the existing methods to extract information chunks from build logs. We then develop prototypical implementations for three promising techniques, namely program synthesis
by example, textual similarity and search keywords. Finally, we evaluate the three techniques in an empirical study on our manually
labeled \emph{LogChunks} data set, which comprises 797 build logs in 29 languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths and weaknesses. We discuss under which circumstances each technique performs best
and provide a recommendation on when developers or researchers should use which
technique.

>>>>

Build logs are a textual by-product that software build processes
create, often as part of their Continuous Integration (CI)
pipeline.
Build logs can serve as a paramount source of
information for developers to understand and debug build or
test failures.
Manually
extracting the important chunks of information, though, is akin to
finding a needle in a haystack.
Recently, researchers and practitioners have begun attempts to partly
automate
this time-consuming activity, with the proposition to not only ease
developers' and researcher's tasks but to enable a new class of automated
on-ward processing applications.
In this paper, we want to give a systematic overview of the state of
the art of the emerging field of build log analysis.
To this end, we first survey and categorize the existing methods to
extract information from build logs.
We then develop prototypical implementations for three promising
techniques, namely program synthesis
by example, textual similarity and search keywords.
Finally, we evaluate the three techniques in an empirical study on
our manually
labeled \emph{LogChunks} data set, which comprises 797 build logs in
29 languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths
and weaknesses.
We discuss under which circumstances each technique performs best
and provide a recommendation on when developers or researchers should
use which technique.

### 2020-04-03 tse-paper/content.tex
small corrections introduction, Closes #20

To gain an impression on how researchers are using information from
build logs
within their research and how they are gathering chunks from them
to support
developers in understanding build outcomes, we conducted a systematic
mapping
study.
Our goal was to classify how and what for researchers use build logs as
a source
of data (\textbf{RQ1}).

@@

To address this, we implement three promising techniques for build
log analysis.
They retrieve specific chunks of targeted information and are configured
by
examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives
(\textbf{RQ2}).

@@

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and targeted log chunks, PBE
  synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by
  the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole
  text for
  the occurrence of specific trigger words.
\end{itemize}

@@

* Try:
  \end{lstlisting}
  \caption{Example of a log chunk explaining why an Android/Gradle
  build failed.}
  % TODO Moritz: we can relate this to travistorrent maybe, you told me
  % long ago that this was something your analyzer cannot extract
  \label{lst:chunk-example-1}
\end{figure}

@@

* Try:
  \end{lstlisting}
  \caption{Another example of a log chunk explaining why an Android/Gradle
  build failed.}
  \label{lst:chunk-example-2}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
StartExtraction(RegexPair("Colon/{\color{Plum}$\circ$}/Line Separator",
"ALL CAPS")) in EndExtraction(RegexPair("/{\color{Plum}$\varepsilon$}/",
"Line Separator/{\color{Plum}$\circ$}/Line Separator"))
  \end{lstlisting}
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example-1,lst:chunk-example-2}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
--- FAIL: TestSeparator (0.03s)
    repo\_test.go:75: expected entry to be in form of `* [link] -
    description.`, got '* ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk.
% TODO: which build tool?
% TODO: move to where structural category explanation
}
  \label{lst:chunk-example-3}
\end{figure}

The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-example-2} show two exemplary
log chunks
from Android/Gradle build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed.
When PBE is trained on these two examples it produces
a regular expression similar to \Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon followed by a line separator and before a capital
letter.
It ends before two line separators, consistent with the two example
chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify the targeted log chunk.
In fact, if the third
example from \Cref{lst:chunk-example-3} with a different structural
representation is added, PBE is not able to syntesize a program and
returns
``\texttt{no program found}''.
We express this difference in structural representation through
dividing log
chunks into structural categories.
The chunks from
\Cref{lst:chunk-example-1,lst:chunk-example-2} are in the same structural
category, while the chunk from \Cref{lst:chunk-example-3} is in a
different one.

@@

Our study of the three techniques on \emph{LogChunks} shows that
\begin{itemize}
  \item PBE yields very accurate results when trained with two examples
  from a
  single structural category.
  \item CTS shows the best average precision, though precision and recall
  of a
  retrieval is hard to determine from the given result.
  A small increase in the number of training examples has no noticeable
  influence.

  Fewer structural categories improve precision and recall of the
  retrieval.
  \item KWS has the highest recall of all techniques, however much
  lower precision.
  It is the technique with the best recall when multiple structural
  categories
  are present in the training examples.
\end{itemize}

@@

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision
and
recall of the chunk retrieval is required.
CTS is well suited when the representation of the desired information
varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired information in
the log
is unpredictable or varies greatly, KWS seems to be the best choice.
However,
its low precision---it extracts a context of multiple lines around a
finding---makes it generally unsuited for automatic on-ward processing and
instead requires a human to further inspect and interpret the output
chunk.

@@

% TODO (Caro): Integrate Petersen's guidelines
From our own experience, we know that the field of analyzing build
logs is relatively young, with perhaps the work on TravisTorrent the
most cited and central one.
We also know there have been scattered
research efforts on build log analysis prior to it, mostly where the
log analysis was a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey them, despite their forseeable future and current importance.

\subsection{Study Selection}
% TODO shorten
\Cref{fig:lit-survey} presents an overview of our study selection.
In this first literature overview study on analyzing build logs, we
therefore want to survey as broad a population of scientific material
as possible.
Note that we do not limit our sources to specific
Software Engineering venues, as is often done, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's theses,
academic slide decks, and books.
We use Google Scholar as opposed to
Scopus or other databases, as it is the search engine for acdemic
articles with the broadest and most current index, including preprints
from sources such as arXiv.
One challenge is that no common
nomenclature has been established yet (in fact, this is one of the
goals of this article).

% TODO Caro cite guidelines
Since we cannot take for granted that ``analyzing build logs'' is the
cannonical term that other, isolated works have used, we instead
decdided to leave out the potentially variable ``analysis'' part and
instead focus on variations of the term ``build log''.
This has the
advantage of being biunique: We explicitly decided to exclude systems
or event logs (see \Cref{sec:rw}).
On the other hand, a Google Scholar
search for ``build log'' returns more than 2.2 million results
(2020-03-15).
We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works.
This left us with 1,552 combined search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.

>>>>

To gain an impression on how researchers are using information from
build logs within their research to support
developers, we conducted a systematic mapping study.
Our goal was to classify how, and what for, researchers use build logs as
a source of data (\textbf{RQ1}).

@@

To address this, we implement three promising techniques for build
log analysis.
They retrieve specific chunks of targeted information and are trained
with examples provided by the user.
We conduct an empirical study on the \emph{LogChunks} data
set~\cite{brandt2020logchunks} to
gauge the performance of these chunk retrieval techniques and characterize
when one of the techniques should be preferred over other alternatives
(\textbf{RQ2}).

@@

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item \textbf{Program Synthesis by Example (referred to as PBE)}
  Based on user examples of build logs and chunks, PBE
  synthesizes
  a regular expression matching the given chunks within the build logs.
  \item \textbf{Common Text Similarity (CTS)}
  Using the Vector Space Model, CTS selects those lines of a log which are
  most similar to the lines present in example chunks provided by
  the user.
  \item \textbf{Keyword Search (KWS)}
  Ad-hoc method for finding fitting passages by searching the whole
  text for
  the occurrence of specific trigger words.
\end{itemize}

@@

* Try:
  \end{lstlisting}
  \caption{Example of a log chunk explaining why an Android
  build failed.}
  % TODO Moritz: we can relate this to travistorrent maybe, you told me
  % long ago that this was something your analyzer cannot extract
  \label{lst:chunk-example-1}
\end{figure}

@@

* Try:
  \end{lstlisting}
  \caption{Another example of a log chunk explaining why an Android
  build failed.}
  \label{lst:chunk-example-2}
\end{figure}

\lstset{language=caml, morekeywords={StartExtraction, RegexPosition,
RegexPair,
  EndExtraction}, keywordstyle=\bfseries\color{black}, escapeinside=//}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
StartExtraction(RegexPair("Colon/$\circ$/Line Separator",
"ALL CAPS")) in EndExtraction(RegexPair("/{\color{Plum}$\varepsilon$}/",
"Line Separator/$\circ$/Line Separator"))
  \end{lstlisting}
  \caption{Simiplified regular expression produced by PBE when trained
  with examples from \Cref{lst:chunk-example-1,lst:chunk-example-2}}
  \label{lst:prose-program-simplified}
\end{figure}

\lstset{
  language=,
  morekeywords={},
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true,frame=tlr]
=== RUN   TestSeparator
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[backgroundcolor=\color{Cerulean!60},breaklines=true,frame=rl]
--- FAIL: TestSeparator (0.03s)
    repo\_test.go:75: expected entry to be in form of `* [link] -
    description.`, got '* ...
  \end{lstlisting}
  \vspace{-\baselineskip}
  \begin{lstlisting}[breaklines=true,frame=blr]
=== RUN   TestGenerateHTML
  \end{lstlisting}
  \caption{Example of a log chunk showing a linter error}
  \label{lst:chunk-example-3}
\end{figure}

The results of our empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techinques.
\Cref{lst:chunk-example-1,lst:chunk-example-2} show two exemplary
log chunks from Android build logs marked in blue.
Both chunks contain the information on why the
corresponding build failed.
When PBE is trained with these two examples it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators, consistent with the two example
chunks from \Cref{lst:chunk-example-1,lst:chunk-example-2}.
This shows, that the characters structuring the text around the chunk are
cruicial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3} with a
different structural representation is added to the training set,
PBE is not able to syntesize a program and returns
``\texttt{no program found}''.
We express this difference in structural representation through
dividing log chunks into structural categories.
The chunks from \Cref{lst:chunk-example-1,lst:chunk-example-2}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a \emph{different structural category}.

@@

% Our study of the three techniques on \emph{LogChunks} shows that
% \begin{itemize}
%   \item PBE yields very accurate results when trained with two examples
%   from a
%   single structural category.
%   \item CTS shows the best average precision, though precision and recall
%   of a
%   retrieval is hard to determine from the given result.
%   A small increase in the number of training examples has no noticeable
%   influence.

%   Fewer structural categories improve precision and recall of the
%   retrieval.
%   \item KWS has the highest recall of all techniques, however much
%   lower precision.
%   It is the technique with the best recall when multiple structural
%   categories
%   are present in the training examples.
% \end{itemize}

@@

Following the results from our comparison study,
recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision
and recall of the chunk retrieval is required.
Results produced by PBE are suited for automatic on-ward processing.
CTS is well suited when the representation of the desired information
varies
slightly and the output of the chunk retrieval is further processed by
a human.
In cases where the textual representation of the desired information in
the log
is unpredictable or varies greatly, KWS seems to be the best choice.
However,
its low precision---it extracts a context of multiple lines around a
finding---makes it generally unsuited for automatic on-ward processing and
instead requires a human to further inspect and interpret the output
chunk.

@@

% TODO (Caro): Integrate Petersen's guidelines
From our own experience, we know that the field of analyzing build
logs is relatively young, with perhaps the work on TravisTorrent the
most cited and central one.
We also know there have been scattered
research efforts on build log analysis, mostly where the
log analysis was a side-part of the article.
To the best of our
knowledge, there has not been an attempt to systematically overview
and survey approaches for processing build logs,
despite their forseeable future and current importance.

\subsection{Study Selection}
% TODO shorten
\Cref{fig:lit-survey} presents an overview of our study selection.
In this first literature overview study on analyzing build logs, we
therefore want to survey as broad a population of scientific material
as possible.
Note that we do not limit our sources to specific
Software Engineering venues, as is often
done~\cite{petersen2015guidelines}, but survey all sources
monitored by Google Scholar, including Bachelor's and Master's theses,
academic slide decks, and books.
We use Google Scholar as opposed to
Scopus or other databases, as it is the search engine for acdemic
articles with the broadest and most current index, including preprints
from sources such as arXiv.
% TODO: maybe remove. this screams snowballing needed
One challenge is that no common
nomenclature has been established yet (in fact, this is one of the
goals of this article).

% TODO Caro cite guidelines
Our Google Scholar
search for variations ``build log'' returns more than 2.2 million results
(2020-03-15).
We therefore refined our search criterions for the four
different spelling possibilities of ``build log'' to exclude obvious
non-related works.
We explicitly decided to exclude systems
or event logs (see \Cref{sec:rw}).
This left us with 1,552 combined search results
(2020-03-15).
The full search queries are documented in our replication
package~\cite{brandt2020chunk-replication}.
